Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-12-06 Documents/Paper/ms.tex
Object --> component

Let $F(x; \hyperparams)$ be a function that takes an object $x$ and transforms it
to a value $u$ that has a uniform distribution between 0 and 1, given $\hyperparams$.
If the object $x$ consists of a single scalar value, $F$ is the cumulative
distribution of the interim prior. Denote the inverse of $F$ by $G$.

A proposal
for an object involves transforming its parameters to $[0, 1]$ using $F$,
making the proposal in terms of $u$, and then transforming back.
Specifically, the proposal chooses
a new value $\xx_i'$ from the current value $\xx_i$ as follows:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
\xx_i' &:=& G(u_i'; \hyperparams).
\end{eqnarray}
where $\delta_u$ is drawn from a heavy-tailed distribution as in
Equation~\ref{eqn:heavytailed}.

Choosing just one object to change is most appropriate when the DNS
distribution is very constrained. When it is close to the prior, bolder
proposals that change more than one object at a time are possible. Hence,
we choose the number of components to change from a heavy tailed distribution
wherer the most probable value is 1 but there is also a nontrivial probability
of proposing to change $\sim N$ components at once.

@@

The ``dragging'' process works as follows, and must be carried out on
each object:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
\xx_i' &:=& G(\xx_i; \hyperparams')
\end{eqnarray}
In other words, the components' parameters are transformed to $[0,1]$ using the
current value of the hyperparameters, and then transformed back using the
proposed value of the hyperparameters so they represent the new interim prior
rather than the old one.

@@

\subsection{Prior Distributions}
To infer $N$, the number of sinusoids, as well as all the periods, amplitudes,
and phases, we need to define prior distributions.
In the notation of Section~\ref{sec:introduction}, the parameters of the
each object are
\begin{eqnarray}
\xx_i &=& \{A_i, T_i, \phi_i\}.
\end{eqnarray}
For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
with mean $\mu$. For the periods, we assigned a log-uniform distribution for
the periods between fixed boundaries, and a uniform distribution for the phases
between 0 and $2\pi$. In other words, our prior is only hierarchical for the
amplitudes $\{A_i\}$. The prior for $\mu$ was a log-uniform prior, with
probability density $\propto 1/\mu$, where $\mu \in [10^{-3}, 10^3]$.

>>>>

Let $F(x; \hyperparams)$ be a function that takes an component $x$ and transforms it
to a value $u$ that has a uniform distribution between 0 and 1, given $\hyperparams$.
If the component $x$ consists of a single scalar value, $F$ is the cumulative
distribution of the interim prior. Denote the inverse of $F$ by $G$.

A proposal
for an component involves transforming its parameters to $[0, 1]$ using $F$,
making the proposal in terms of $u$, and then transforming back.
Specifically, the proposal chooses
a new value $\xx_i'$ from the current value $\xx_i$ as follows:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
\xx_i' &:=& G(u_i'; \hyperparams).
\end{eqnarray}
where $\delta_u$ is drawn from a heavy-tailed distribution as in
Equation~\ref{eqn:heavytailed}.

Choosing just one component to change is most appropriate when the DNS
distribution is very constrained. When it is close to the prior, bolder
proposals that change more than one component at a time are possible. Hence,
we choose the number of components to change from a heavy tailed distribution
wherer the most probable value is 1 but there is also a nontrivial probability
of proposing to change $\sim N$ components at once.

@@

The ``dragging'' process works as follows, and must be carried out on
each component:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
\xx_i' &:=& G(\xx_i; \hyperparams')
\end{eqnarray}
In other words, the components' parameters are transformed to $[0,1]$ using the
current value of the hyperparameters, and then transformed back using the
proposed value of the hyperparameters so they represent the new interim prior
rather than the old one.

@@

\subsection{Prior Distributions}
To infer $N$, the number of sinusoids, as well as all the periods, amplitudes,
and phases, we need to define prior distributions.
In the notation of Section~\ref{sec:introduction}, the parameters of the
each component are
\begin{eqnarray}
\xx_i &=& \{A_i, T_i, \phi_i\}.
\end{eqnarray}
For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
with mean $\mu$. For the periods, we assigned a log-uniform distribution for
the periods between fixed boundaries, and a uniform distribution for the phases
between 0 and $2\pi$. In other words, our prior is only hierarchical for the
amplitudes $\{A_i\}$. The prior for $\mu$ was a log-uniform prior, with
probability density $\propto 1/\mu$, where $\mu \in [10^{-3}, 10^3]$.
